---
title: "Creating a presence absence classifier for Pennatula phosphorea"
author: "James Hutchison"
date: "21 March 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The first stage is to set up folders containing images for training, validation and testing for each output class from the model. We have two output classes (present or absent), so will need 6 folders. From Marine Recorder we have 1161 images with Pennatula phosphorea present and 1003 images from the same set of surveys where it is absent. The presence and absence images are stored in two directories. We will use R to copy 600 to a training directory, 200 to a validation directory and the rest to a testing directory for each class.

```{r create directories}
base_dir <- "D:/Image analysis/PenPhosMR/PenPhos"

train_dir <- file.path(base_dir, "train")
dir.create(train_dir)

validation_dir <- file.path(base_dir, "validation")
dir.create(validation_dir)

test_dir <- file.path(base_dir, "test")
dir.create(test_dir)

train_present_dir <- file.path(train_dir, "present")
dir.create(train_present_dir)

train_absent_dir <- file.path(train_dir, "absent")
dir.create(train_absent_dir)

validation_present_dir <- file.path(validation_dir, "present")
dir.create(validation_present_dir)

validation_absent_dir <- file.path(validation_dir, "absent")
dir.create(validation_absent_dir)

test_present_dir <- file.path(test_dir, "present")
dir.create(test_present_dir)

test_absent_dir <- file.path(test_dir, "absent")
dir.create(test_absent_dir)

presence_files <- list.files(file.path(base_dir, "Present_best"))
presence_train <- sample(1:length(presence_files), round(0.6*length(presence_files)))
presence_validation <- sample((1:length(presence_files))[-presence_train], round(0.2*length(presence_files)))
presence_test <- (1:length(presence_files))[-c(presence_train, presence_validation)]

file.copy(file.path(base_dir, "Present", presence_files[presence_train]), train_present_dir)
file.copy(file.path(base_dir, "Present", presence_files[presence_validation]), validation_present_dir)
file.copy(file.path(base_dir, "Present", presence_files[presence_test]), test_present_dir)

absence_files <- list.files(file.path(base_dir, "Not present"))
absence_train <- sample(1:length(absence_files), round(0.6*length(absence_files)))
absence_validation <- sample((1:length(absence_files))[-absence_train], round(0.2*length(absence_files)))
absence_test <- (1:length(absence_files))[-c(absence_train, absence_validation)]

file.copy(file.path(base_dir, "Not present", absence_files[absence_train]), train_absent_dir)
file.copy(file.path(base_dir, "Not present", absence_files[absence_validation]), validation_absent_dir)
file.copy(file.path(base_dir, "Not present", absence_files[absence_test]), test_absent_dir)
```


```{r train model}
library(keras)

#conv_base <- application_vgg16(weights = "imagenet", include_top=FALSE, input_shape = c(150,150,3))
conv_base <- application_inception_resnet_v2(weights = "imagenet", include_top=FALSE, input_shape = c(299,299,3))

model <- keras_model_sequential() %>% 
  conv_base %>% 
  layer_flatten() %>% 
  layer_dense(units = 256, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")

#train_datagen = image_data_generator(
#  rescale = 1/255,
#  rotation_range = 40,
#  width_shift_range = 0.2,
#  height_shift_range = 0.2,
#  shear_range = 0.2,
#  zoom_range = 0.2,
#  horizontal_flip = TRUE,
#  fill_mode = "nearest"
#)

#test_datagen <- image_data_generator(rescale = 1/255)

train_generator <- flow_images_from_directory(
  train_dir,                  # Target directory  
  #train_datagen,              # Data generator
  target_size = c(299, 299),  # Resizes all images to 150 × 150
  batch_size = 20,
  class_mode = "binary"       # binary_crossentropy loss for binary labels
)

validation_generator <- flow_images_from_directory(
  validation_dir,
  #test_datagen,
  target_size = c(299, 299),
  batch_size = 20,
  class_mode = "binary"
)

model %>% compile(
  loss = "binary_crossentropy",
  optimizer = optimizer_rmsprop(lr = 2e-5),
  metrics = c("accuracy")
)

history <- model %>% fit_generator(
  train_generator,
  steps_per_epoch = 100,
  epochs = 30,
  validation_data = validation_generator,
  validation_steps = 50
)

```

```{r test}

test_generator <- flow_images_from_directory(
  test_dir,
  target_size = c(299, 299),
  batch_size = 20,
  class_mode = "binary",
  shuffle = FALSE
)

model %>% evaluate_generator(test_generator, steps = 50)


```