---
title: "Flickr Image Scraper"
author: "Mark Logie"
date: "16 April 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This is a script for scraping photos from Flickr.
Before running this script, you need to create a csv file with all your search terms.  This csv needs to be structured as follows:

Target             | SearchTerm1        | SearchTerm2         | SearchTerm3     | ...
------------------ | ------------------ | ------------------- | --------------- | ---
Alsike Clover      | Alsike Clover      | Trifolium hybridum  |                 | ...
Birds Foot Trefoil | Birds Foot Trefoil | Lotus corniculatus  |                 | ...
Black Knapweed     | Black Knapweed     | Centaurea debeauxii | Centaurea nigra | ...
Black Medick       | Black Medick       | Medicago lupulina   |                 | ...
Lucerne            | Lucerne Flower     | Medicago sativa     | Alfalfa Flower  | ...

* The first row can be anything you want.
* The first column contains all the search targets: a list of everything you would like to find images of, but not necessarily the search terms.
* The next columns are the actual search terms, which will be combined into folders for each target.

Once you've created this csv, save it somewhere and set that location here:
```{r csv}
pathtocsvs <- file.path('.','Image_Retrieval','Flickr')
```

## Dependencies

Here are the dependencies for the script

```{r dependencies, message=FALSE, warning=FALSE}
source(file.path('.','Image_Retrieval','Flickr','FlickR.R'))
library(devtools)
install_github('FrancescaMancini/FlickrAPI_EABhackathon')
library(flickr)
library(httr)
library(RCurl)
```

## Setting up initial parameters

This section of the script sets up the parameters for the search and where the photos are going to be saved.
The parameters to be set up are:

* Save directory for the images once downloaded
* The number of images to be saved for each target for example:
    + maximages = 100
    + maximages = Inf
* The start and finish years for the search
    + Either of these can be set to the current year using 'format(Sys.Date(), "%Y")'
* Finally, the location in the world where you want to search for images.
    + Good options are Scotland, UK, Europe, or World
    + If instead you want to search everywhere, set this to NULL

```{r parameters}
savelocation <- file.path('.','Image_Retrieval','Flickr')
maximages <- Inf
startYear <- 1990
endYear   <- format(Sys.Date(), "%Y")
worldlocation <- NULL
```
## Flickr Authentication

You need to authenticate with Flickr, to allow you to query the Flickr database.  The below command needs to be run once.
It is currently set to not be evaluated.  If this is your first time running this script, you need to run this command in the console.

```{r flickr_authentication, eval=FALSE}
authFlickr()
```
## Downloader

Now that you're all set up, the script will run and download all the images for the search terms you've specified.  This can take a long time.

First, the script find the relevant ID number for the location you chose (if you chose a location), and outputs the name.
Then the images are downloaded.

```{r downloader}
if(!is.null(worldlocation)){
  woeResult <- findPlaces(worldlocation)
  cat(paste0('Location chosen: ',woeResult$woe_name[1],'\n'))
}
csvs <- list.files(pathtocsvs,'\\.csv')

FileError <- c()
if(length(csvs)==0){
  cat('This script requires a search csv. Please create one and try again\n')
} else if(length(csvs)>1){
  cat('This script requires exactly 1 search csv. Please delete extra csvs and try again\n')
} else {
  searchDF <- read.csv(file.path(pathtocsvs,csvs),stringsAsFactors = FALSE)
  download.flickr(searchDF,savelocation,maximages,
                  woeResult = NULL,c(startYear,endYear),FileError)
  cat(FileError)
}
```